### 2021-04-20 input/figproof.tex
issue depiction
purging

\begin{figure*}
  \begin{center}
    \subfloat[Part A][\label{fig:proofA}Stale $\alpha$'s may stop
      other $\alpha$'s from reaching all processes ($b$ and $c$) that
      require it along the shortest path. Hence, we must guarantee
      eventual purging of stale $\alpha$'s (see
      Figure~\ref{fig:proofB}).]  {\input{input/figproofA.tex}}
    \hspace{10pt}
    \subfloat[Part B][\label{fig:proofB}Stale $\alpha$'s may
      stop $\delta$'s from reaching
      processes with targeted $\alpha$'s.
      To ensure consistent partitioning, $b'$ and $b''$
      must remove $\alpha_{d}$ either by delivering $\delta_d$,
      or a more specific $\delta_d^{0.5}$ and $\delta_d^{1.5}$ respectively.]
    {\input{input/figproofB.tex}}
    \caption{\label{fig:proof}Dynamic partitioning leads to correctness issues due to
      staleness and ordering of operations.}
  \end{center}
\end{figure*}

>>>>

\begin{figure*}
  \begin{center}
    \subfloat[Part A][\label{fig:proofA}Stale $\alpha$ messages may
      stop other $\alpha$ messages from reaching all \processes ($b$
      and $c$) that require it along the shortest path. Consistent
      partitioning requires eventual purging of stale messages (see
      Figure~\ref{fig:proofB}).]  {\input{input/figproofA.tex}}
    \hspace{10pt}
    \subfloat[Part B][\label{fig:proofB}Stale $\alpha$ messages may
      stop $\delta$ messages from reaching processes with the corresponding
      partition. Consistent partitioning requires that all \processes propagate
      $\delta$ messages as often as possible (I'); detect possible inconsistencies
      when parents' $\alpha$ messages contradict delivery history (III); notify
      processes of possible inconsistencies (IV).]
    {\input{input/figproofB.tex}}
    \caption{\label{fig:proof}Concurrent operations,
      delivery orders, and stale control information may impair
      consistent partitioning. Dynamic consistent partitioning
      requires eventual purging of stale messages, and best eventual
      forwarding.}
  \end{center}
\end{figure*}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../paper"
%%% ispell-local-dictionary: "english"
%%% End:

### 2021-04-20 input/proposal.tex
purging

At any time, a \process can decide to become a source, hence creating a
new partition in the system by executing an \texttt{Add}
operation. This partition includes at least its source plus
neighboring \processes that estimate they are closer to this source
than any other source. Such distance (or \emph{weight}) is
application-dependant: in the context of maintaining distributed
indexes, weights are latency.

%For instance, weights could be latency in live
% streaming applications, or differences in profiles for recommendation
%systems.

@@

\begin{corollary}[\label{theo:dcp}\underline{D}ynamic \underline{c}onsistent
    \underline{p}artitioning (DCP)]
  Consistent partitioning where each \process can \texttt{Add} and
  \texttt{Del}ete its partition at any time requires
  \begin{inparaenum}[(i)]
  \item eventual purging of stale notifications, and
  \item eventual delivery of bests partitions.
  \end{inparaenum}
\end{corollary}

\begin{proof}
  \Processes deliver and forward the best sum of weights they
  received. Figure~\ref{fig:proofA} depicts the only case where a
  \process would block such epidemic propagation: the \process
  received and delivered a message from a partition that has since
  then been deleted but with a better sum of weights than newly
  received ones. Removing \emph{forever} all such stale messages about
  deleted partitions would allow \processes to propagate their best
  up-to-date partition again, eventually reaching consistent
  partitioning together, as stated by Theorem~\ref{theo:bef}.
\end{proof}

@@

Figure~\ref{fig:proof} depicts the issues with staleness and message
orderings. In Figure~\ref{fig:proofA}, the shortest path from any
source to \Process $c$ is $[a, b, c]$. However, \Process $b$ still holds
a stale $\alpha_d^{0.5}$ without knowing. When it receives
$\alpha_a^1$, it discards it, for it assumes that downstream \processes
are more interested in $\alpha_d^{0.5}$. To reach consistent
partitioning, \Process $b$ first needs to purge its current partition
to later accept that of its current actual shortest path:
$\alpha_a^1$.

\noindent Figure~\ref{fig:proofB} shows that removing stale control
information is even more complex. The removal must reach all \processes
of the previous shortest path going from $d$ to $b''$. Label I' shows
the most obvious issue where \Process $e$ changed partition for a
better but stale $\alpha_f$. Since it can remember its previous
deliveries, it could still forward $\delta_d$ for the sake of
consistency. However, this would lead to every \process forwarding
every $\delta$ they ever delivered. Such protocol's overhead would
depend on past partitioning instead of current one. Label III shows
the issue when the blocking partition is already known to be stale at
\Process $b'$. \Process $b''$ eventually receives $\alpha_f$ from
$e$. However, it cannot deliver it, for it would break last win order.
\Process $b'$ may be in an inconsistent state. Label IV shows the
corollary issue: \Process $b''$ delivered a message from a \process that
may be inconsistent without knowing it.

\begin{theorem}[Propagation $+$ detection $\implies$ DCP]
  % Propagating 
  \TODO{Puring stale notifications requires to handle I...IV.}
\end{theorem}

\begin{proof}
  \TODO{Meow meow meow.}
  \TODO{Label I: default propagation to what seems the border of the partition but isn't.}
  \TODO{Label II: gates eventually break because of I', III, IV.}
  \TODO{Label I': as Label I.}
  \TODO{Label III: detecting possible inconsistent partitioning.}
  \TODO{Label IV: propagation of this issue to all and only concerned parties.}
\end{proof}

>>>>

At any time, a \process can decide to become a source, hence creating
a new partition in the system by executing an \texttt{Add}
operation. This partition includes at least its source plus
neighboring \processes that estimate they are closer to this source
than any other source. Such distance (or \emph{weight}) is
application-dependant: in the context of maintaining distributed
indexes, weights would be link latency.

@@

\begin{corollary}[\label{theo:dcp}purge + BEF $\implies$
    \underline{D}ynamic \underline{c}onsistent
    \underline{p}artitioning (DCP)]
%
When each \process can \texttt{Add} and \texttt{Del}ete its partition
at any time, consistent partitioning requires
\begin{inparaenum}[(i)]
\item eventual purging of stale messages ($b_p(m) \rightarrow b_p(m')
  \implies \eventually (d_q(m) \implies d_q(m) \rightarrow
  d_q(m''))$);
\item and best eventual forwarding.
\end{inparaenum}
\end{corollary}

\begin{proof}
  \Processes deliver and forward the best sum of weights they
  received. Figure~\ref{fig:proofA} depicts the only case where a
  \process would block such epidemic propagation: the \process
  received and delivered a message from a partition that has since
  then been deleted but with a better sum of weights than newly
  received ones (in the example $\alpha_d^y <
  \alpha_a^{x'}$). Removing \emph{forever} all such stale messages
  about deleted partitions would allow \processes to propagate their
  best up-to-date partition again, eventually reaching consistent
  partitioning together, as stated by Theorem~\ref{theo:bef}.
\end{proof}

@@

%% Figure~\ref{fig:proof} depicts the issues with staleness and message
%% orderings. In Figure~\ref{fig:proofA}, the shortest path from any
%% source to \Process $c$ is $[a, b, c]$. However, \Process $b$ still holds
%% a stale $\alpha_d^{0.5}$ without knowing. When it receives
%% $\alpha_a^1$, it discards it, for it assumes that downstream \processes
%% are more interested in $\alpha_d^{0.5}$. To reach consistent
%% partitioning, \Process $b$ first needs to purge its current partition
%% to later accept that of its current actual shortest path:
%% $\alpha_a^1$.

%% \noindent Figure~\ref{fig:proofB} shows that removing stale control
%% information is even more complex. The removal must reach all \processes
%% of the previous shortest path going from $d$ to $b''$. Label I' shows
%% the most obvious issue where \Process $e$ changed partition for a
%% better but stale $\alpha_f$. Since it can remember its previous
%% deliveries, it could still forward $\delta_d$ for the sake of
%% consistency. However, this would lead to every \process forwarding
%% every $\delta$ they ever delivered. Such protocol's overhead would
%% depend on past partitioning instead of current one. Label III shows
%% the issue when the blocking partition is already known to be stale at
%% \Process $b'$. \Process $b''$ eventually receives $\alpha_f$ from
%% $e$. However, it cannot deliver it, for it would break last win order.
%% \Process $b'$ may be in an inconsistent state. Label IV shows the
%% corollary issue: \Process $b''$ delivered a message from a \process that
%% may be inconsistent without knowing it.

\begin{theorem}[Propagations $+$ detection $\implies$ purge]
  \Processes can purge stale control information by broadcasting a
  dedicated notification message $\delta$, and detect possible
  inconsistencies using their history to propagate 
\end{theorem}

\begin{proof}
  Figure~\ref{fig:proofB} depicts all possible cases that need
  appropriate purging. A chain of \processes $[d, e, b', b'']$
  delivered and forwarded a message $\alpha_d$. Some \processes may
  have delivered messages about other partitions before or after
  $\alpha_d$.
  \begin{asparadesc}
  \item [\processes with last $\alpha_d$:] \Processes such as
    \Process~$d$ where the last and best partition is still $P_d$. By
    propagating from process to process the corresponding $\delta_d$
    message, these \processes purge $\alpha_d$.
  \item [\processes with $\alpha_d^y \rightarrow \alpha_f^z$ with last
    $\alpha_f^z$, for $\alpha_f^z < \alpha_d^y$:] \Processes such as
    \Process~$e$ do not directly suffer from the non-delivery of
    $\delta_d$. They need to purge their own $P_f$ if need
    be. Nonetheless, they deliver and forward $\alpha_f$ that at least
    one following \process will receive:
  \item [\processes with $\delta_f \rightarrow \alpha_d$:] \Processes
    such as \Process~$b'$ do not belong to the preceding category, for
    they already delivered $\delta_f$. However, best eventual
    forwarding guarantees that they receive $\alpha_f$ that
    contradicts their history or state. In such case, the detecting
    \processes purge their own $\alpha_d$. The detecting \processes
    must also notify subsequent \processes that their current
    partition \emph{may be} deleted, for their parent may have blocked
    the corresponding $\delta$ messages. We note such notification
    $\delta_d?$ to emphasize the uncertainty of a $\delta$ message.
  \item [\processes with last $\alpha_d$ receiving $\delta_d?$ from
    their parent:] \Processes such as \Process~$b''$ have $\alpha_d$
    and preceding delivered messages have not importance. Such
    \processes must trust detecting \processes by delivering and
    forwarding $\delta_d?$. Messages are subject to all aforementioned
    blocking conditions, but are also solved by aforementioned
    mechanisms.
  \end{asparadesc}

  Propagation I' terminates: a process do not deliver a $\delta$
  message if it has already delivered it. Propagation IV terminates: a
  process do not deliver a looping message.  Propagation and detection
  ensure that \processes eventually purge stale $\alpha$ messages from
  the system, leaving room for Propagation I.
  %% In the meantime, \processes that still belong to a partition can
  %% propagate their respective last $\alpha$ message, and reach
  %% consistent partitioning.
\end{proof}
